---
title: 'TCR Analysis'
author: |
 | Derrik Gratz
 | FHIL
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
    fig_width: 9
    toc_depth: 3
    css: "robobook.css"
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE,
                      fig.width = 6, fig.height = 6, dev='svglite',
                      cache=FALSE) 
library(kableExtra)
```

# Overview

This report shows data from the bulk TCR sequencing from the BM03 project processed using [MixCR](https://mixcr.com/) and analyzed with [Immunarch](https://immunarch.com/). Due to differences in sequencing depth and saturation across kits and samples, **all data shown in this report is downsampled to the lowest UMI count of any sample for that RNA input (L/S)**. Large input samples were downsampled to 311422 UMIs. Small input samples were downsampled to 9961 UMIs.

This report includes data from all sequencing done as of Aug 21 2024. Data is pulled from these repositories:

Takara:

`/fh/fast/_IRC/FHIL/grp/NextSeq_SteamPlant/240723_VH00738_256_AAFWG5MM5/FHIL_07232024_2024-07-25T19_16_58_f111bc7-425851455/BCL_Convert_07_25_2024_12_22_52-752495789/`

celecta:

`/fh/fast/_IRC/FHIL/grp/NextSeq_SteamPlant/240813_VH00738_265_AAG3G77M5/FHIL_08132024_2024-08-14T21_59_11_2bb08af-428286878/BCL_Convert_08_14_2024_15_02_05-760993109/`

neb:

`/fh/fast/_IRC/FHIL/grp/NextSeq_SteamPlant/240730_VH00738_259_AAFWGHHM5/FHIL_07302024_2024-07-30T17_34_21_8f33e36-426337156/BCLConvert_08_01_2024_00_22_32Z-755097350/`

qiaseq:

`/fh/fast/_IRC/FHIL/grp/NextSeq_SteamPlant/240614_VH00738_246_AAFVY2MM5/FHIL_06142024_2024-06-14T19_50_58_68f53f1-421632874/BCLConvert_06_15_2024_16_53_29Z-740072376/`

***

<!-- # Clone recovery -->

<!-- Number of recovered chains/clones -->

<!-- ## Large {.tabset} -->

<!-- ### TRA/TRB -->

<!-- ```{r, fig.width = 6, fig.height = 4} -->
<!-- figures[['chain_recovery_large_TRATRB']] -->
<!-- ``` -->

<!-- ### Other -->

<!-- ```{r} -->
<!-- figures[['chain_recovery_large_other']] -->
<!-- ``` -->


<!-- ## Small {.tabset} -->

<!-- ### TRA/TRB -->

<!-- ```{r, fig.width = 6, fig.height = 4} -->
<!-- figures[['chain_recovery_small_TRATRB']] -->
<!-- ``` -->

<!-- ### Other -->

<!-- ```{r, fig.width = 6, fig.height = 4} -->
<!-- figures[['chain_recovery_small_other']] -->
<!-- ``` -->

<!-- *** -->

# Unique clone count {.tabset}

Number of unique chains/clones

## Large {.tabset .unlisted}

### TRATRB

```{r, fig.width = 6, fig.height = 4}
figures[['unique_clones_large_TRATRB']]
```

### other

```{r, fig.width = 6, fig.height = 4}
figures[['unique_clones_large_other']]
```

## Small {.tabset .unlisted}

### TRATRB

```{r, fig.width = 6, fig.height = 4}
figures[['unique_clones_small_TRATRB']]
```

### other

```{r, fig.width = 6, fig.height = 4}
figures[['unique_clones_small_other']]
```

# {- .unlisted}

***

# Clone count proportions {.tabset}

Relative abundance for the groups of rare clonotypes with low counts

## Large {.unlisted}

```{r, fig.width = 8, fig.height = 4}
figures[['clone_count_proportion_large_TRATRB']]
```

## Small {.unlisted}

```{r, fig.width = 8, fig.height = 4}
figures[['clone_count_proportion_small_TRATRB']]
```

***

<!-- # Clone frequency proportion -->

<!-- Proportion of repertoire occupied by clonal groups with specific abundances. -->

<!-- ## Large -->

<!-- ```{r, fig.width = 8, fig.height = 4} -->
<!-- figures[['clone_frequency_proportion_large_TRATRB']] -->
<!-- ``` -->

<!-- ## Small -->

<!-- ```{r, fig.width = 8, fig.height = 4} -->
<!-- figures[['clone_frequency_proportion_small_TRATRB']] -->
<!-- ``` -->

<!-- *** -->

# {- .unlisted}

# Diversity estimation {.tabset}

Inverse Simpson index - effective number of types that is obtained when the weighted arithmetic mean is used to quantify average proportional abundance of types in the dataset of interest.

## Large {.unlisted}

```{r, fig.width = 6, fig.height = 4}
figures[['simpson_large']]
```

## Small {.unlisted}

```{r, fig.width = 6, fig.height = 4}
figures[['simpson_small']]
```

***


# {- .unlisted}

# Gene usage {.tabset}

Measuring similarity in gene usage across samples

## Large {.tabset .unlisted}

### TRAJ

```{r, fig.width=8, fig.height=7}
figures[['traj_usage_cor_large']]
```

### TRAV

```{r, fig.width=8, fig.height=7}
figures[['trav_usage_cor_large']]
```

### TRBJ

```{r, fig.width=8, fig.height=7}
figures[['trbj_usage_cor_large']]
```

### TRBV

```{r, fig.width=8, fig.height=7}
figures[['trbv_usage_cor_large']]
```

## Small {.tabset .unlisted}

### TRAJ

```{r, fig.width=8, fig.height=7}
figures[['traj_usage_cor_small']]
```

### TRAV

```{r, fig.width=8, fig.height=7}
figures[['trav_usage_cor_small']]
```

### TRBJ

```{r, fig.width=8, fig.height=7}
figures[['trbj_usage_cor_small']]
```

### TRBV

```{r, fig.width=8, fig.height=7}
figures[['trbv_usage_cor_small']]
```

# {- .unlisted}

***

# Clonal overlap 

## Heatmaps {.tabset}

Measuring similarity between samples by overlap of repertoires. I did a spot check looking at the top portion of the repertoire rather than the full rep but results looked mostly the same, so this still uses full repertoire. 

### All {.tabset}

#### TRA

```{r, fig.width=8, fig.height=7}
figures[['overlap_all_tra']] 
```

#### TRB

```{r, fig.width=8, fig.height=7}
figures[['overlap_all_trb']] 
```

### Small {.tabset}

#### TRA

```{r, fig.width=8, fig.height=7}
figures[['overlap_small_tra']] 
```

#### TRB

```{r, fig.width=8, fig.height=7}
figures[['overlap_small_trb']] 
```

### Large {.tabset}

#### TRA

```{r, fig.width=8, fig.height=7}
figures[['overlap_large_tra']] 
```

#### TRB

```{r, fig.width=8, fig.height=7}
figures[['overlap_large_trb']] 
```

***


## MDS {.tabset}

Multiple dimension scaling of clonal overlap.

### Small {.tabset}

#### TRB

```{r}
figures[['clonal_overlap_mds_small_TRB']]
```

#### TRA

```{r}
figures[['clonal_overlap_mds_small_TRA']]
```

### Large {.tabset}

#### TRB

```{r}
figures[['clonal_overlap_mds_large_TRB']]
```

#### TRA

```{r}
figures[['clonal_overlap_mds_large_TRA']]
```

***

# Shared clone correlation {.tabset}

## All

```{r, fig.height=16, fig.width=16}
figures[['spearman_cor_all']]
```


## F2

```{r, fig.height=6, fig.width=6}
figures[['spearman_cor_F2']]
```

## F1

```{r, fig.height=6, fig.width=6}
figures[['spearman_cor_F1']]
```

***

# Jaccard {.tabset}

## All

```{r, fig.height=16, fig.width=16}
figures[['jaccard_all_tra']]
```


## F2

```{r, fig.height=6, fig.width=6}
figures[['jaccard_F2_tra']] 
```

## F1

```{r, fig.height=6, fig.width=6}
figures[['jaccard_cor_F1']]
```

