---
title: 'TCR Analysis'
author: |
 | Derrik Gratz
 | FHIL
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
    fig_width: 9
    toc_depth: 3
    css: "robobook.css"
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE,
                      fig.width = 6, fig.height = 6, dev='svglite',
                      cache=FALSE) 
library(kableExtra)
```

# Overview

This report shows data from the bulk TCR sequencing from the BM03 project processed using [MixCR](https://mixcr.com/) and analyzed with [Immunarch](https://immunarch.com/). Due to differences in sequencing depth and saturation across kits and samples, **all data shown in this report is downsampled to the lowest UMI count of any sample for that RNA input (L/S)**. Large input samples were downsampled to 311422 UMIs. Small input samples were downsampled to 9961 UMIs.

Samples follow a naming convention describing 

* The kit used
* The source sample ID
* The input RNA amount (L=large, s=small)
* The replicate ID (if applicable, not every sample had replicates)

***

<!-- # Clone recovery -->

<!-- Number of recovered chains/clones -->

<!-- ## Large {.tabset} -->

<!-- ### TRA/TRB -->

<!-- ```{r, fig.width = 6, fig.height = 4} -->
<!-- figures[['chain_recovery_large_TRATRB']] -->
<!-- ``` -->

<!-- ### Other -->

<!-- ```{r} -->
<!-- figures[['chain_recovery_large_other']] -->
<!-- ``` -->


<!-- ## Small {.tabset} -->

<!-- ### TRA/TRB -->

<!-- ```{r, fig.width = 6, fig.height = 4} -->
<!-- figures[['chain_recovery_small_TRATRB']] -->
<!-- ``` -->

<!-- ### Other -->

<!-- ```{r, fig.width = 6, fig.height = 4} -->
<!-- figures[['chain_recovery_small_other']] -->
<!-- ``` -->

<!-- *** -->

# Unique clone count {.tabset}

Number of unique chains/clones

## Large {.tabset .unlisted}

### TRATRB

```{r, fig.width = 6, fig.height = 4}
figures[['unique_clones_large_TRATRB']]
```

### other

```{r, fig.width = 6, fig.height = 4}
figures[['unique_clones_large_other']]
```

## Small {.tabset .unlisted}

### TRATRB

```{r, fig.width = 6, fig.height = 4}
figures[['unique_clones_small_TRATRB']]
```

### other

```{r, fig.width = 6, fig.height = 4}
figures[['unique_clones_small_other']]
```

# {- .unlisted}

***

# Clone count proportions {.tabset}

Relative abundance for the groups of rare clonotypes with low counts

## Large {.unlisted}

```{r, fig.width = 8, fig.height = 4}
figures[['clone_count_proportion_large_TRATRB']]
```

## Small {.unlisted}

```{r, fig.width = 8, fig.height = 4}
figures[['clone_count_proportion_small_TRATRB']]
```

***

<!-- # Clone frequency proportion -->

<!-- Proportion of repertoire occupied by clonal groups with specific abundances. -->

<!-- ## Large -->

<!-- ```{r, fig.width = 8, fig.height = 4} -->
<!-- figures[['clone_frequency_proportion_large_TRATRB']] -->
<!-- ``` -->

<!-- ## Small -->

<!-- ```{r, fig.width = 8, fig.height = 4} -->
<!-- figures[['clone_frequency_proportion_small_TRATRB']] -->
<!-- ``` -->

<!-- *** -->

# {- .unlisted}

# Diversity estimation {.tabset}

Inverse Simpson index - effective number of types that is obtained when the weighted arithmetic mean is used to quantify average proportional abundance of types in the dataset of interest.

## Large {.unlisted}

```{r, fig.width = 6, fig.height = 4}
figures[['simpson_large']]
```

## Small {.unlisted}

```{r, fig.width = 6, fig.height = 4}
figures[['simpson_small']]
```

***


# {- .unlisted}

# Gene usage {.tabset}

Measuring similarity in gene usage across samples

## Large {.tabset .unlisted}

### TRAJ

```{r, fig.width=8, fig.height=7}
figures[['traj_usage_cor_large']]
```

### TRAV

```{r, fig.width=8, fig.height=7}
figures[['trav_usage_cor_large']]
```

### TRBJ

```{r, fig.width=8, fig.height=7}
figures[['trbj_usage_cor_large']]
```

### TRBV

```{r, fig.width=8, fig.height=7}
figures[['trbv_usage_cor_large']]
```

## Small {.tabset .unlisted}

### TRAJ

```{r, fig.width=8, fig.height=7}
figures[['traj_usage_cor_small']]
```

### TRAV

```{r, fig.width=8, fig.height=7}
figures[['trav_usage_cor_small']]
```

### TRBJ

```{r, fig.width=8, fig.height=7}
figures[['trbj_usage_cor_small']]
```

### TRBV

```{r, fig.width=8, fig.height=7}
figures[['trbv_usage_cor_small']]
```

# {- .unlisted}

***

# Clonal overlap 

## Heatmaps {.tabset}

Measuring similarity between samples by overlap of repertoires. I did a spot check looking at the top portion of the repertoire rather than the full rep but results looked mostly the same, so this still uses full repertoire. 

### All {.tabset}

#### TRA

```{r, fig.width=8, fig.height=7}
figures[['overlap_all_tra']] 
```

#### TRB

```{r, fig.width=8, fig.height=7}
figures[['overlap_all_trb']] 
```

### Small {.tabset}

#### TRA

```{r, fig.width=8, fig.height=7}
figures[['overlap_small_tra']] 
```

#### TRB

```{r, fig.width=8, fig.height=7}
figures[['overlap_small_trb']] 
```

### Large {.tabset}

#### TRA

```{r, fig.width=8, fig.height=7}
figures[['overlap_large_tra']] 
```

#### TRB

```{r, fig.width=8, fig.height=7}
figures[['overlap_large_trb']] 
```

***


## MDS {.tabset}

Multiple dimension scaling of clonal overlap.

### Small {.tabset}

#### TRB

```{r}
figures[['clonal_overlap_mds_small_TRB']]
```

#### TRA

```{r}
figures[['clonal_overlap_mds_small_TRA']]
```

### Large {.tabset}

#### TRB

```{r}
figures[['clonal_overlap_mds_large_TRB']]
```

#### TRA

```{r}
figures[['clonal_overlap_mds_large_TRA']]
```